Spatial phylogenetics of Japanese ferns: Patterns, processes, and implications for conservation

### 2022-01-21 ms/SI.Rmd
Update MS
Fix dendrogram in SI where non-monophyletic 'other' was causing large portion of tree to be colored grey

```{r load-data-si, cache = FALSE}
tar_load(c(
  "japan_shp",
  "inext_res",
  "redundancy_by_res",
  "redundancy_summary",
  "biodiv_ferns_spatial",
  "biodiv_ferns_endemic_spatial",
  "k_taxonomy",
  "k_phylogeny",
  "traits_nmds",
  "ppgi",
  "phy_sig_results",
  "binary_sig_results",
  "lrt_results",
  "japan_shp",
  "morans_i",
  "t_test_results",
  "aic_env_repro",
  "comm_ferns",
  "repro_data",
  "japan_pterido_tree",
  "model_fits_not_ult",
  "biodiv_ferns_cent",
  "protected_areas",
  "deer_range",
  "biodiv_ferns_spatial_lumped",
  "bioregion_cutoff",
  "ms_functions"
))
```

@@

**`r s_figure("grain-size")`**. Effect of grain size on sampling redundancy. (A), Percent of grid cells with redundancy > 0.5 by grain size (side length of square grid cells set to 10 km, 20 km, 30 km, or 40 km). (B), Percent of grid cells with redundancy = 0 by grain size. (C), Distribution of redundancy values by grain size.

@@

**`r s_figure("envir")`**. Reproductive mode and environmental data in 20 km × 20 km grid cells across Japan. (A), rolling mean area (km^2^) in 20 km latitudinal bands.  (B), % apomictic taxa. (C), annual precipitation. (D), percipitation seasonality (coefficient of variation). (E), mean annual temperature. (F), temperature seasonality (standard deviation × 100). Data excluded from models not shown in this and subsequent plots unless otherwise mentioned.

@@

**`r s_figure("abundance")`**. Observed number of specimens (A) and sampling redundancy (B) per 20 km grid cell in the ferns of Japan.

@@

**`r s_figure("raw-lat")`**. (A) Raw taxonomic diversity (B), phylogenetic diversity, and (C) functional diversity of the ferns of Japan plotted by latitude. For (B) and (C), raw branchlengths were transformed to relative values by dividing each branchlength by the sum of all branchlengths. Trendlines fit with general additive model using `r func("geom_smooth")` function in the `r pack("ggplot2")` R package [@Wickham2016].

@@

**`r s_figure("raw-biplot")`**. Relationships between observed functional and phylogenetic diversity and taxonomic richness in the ferns of Japan. (A), Phylogenetic diversity (PD). (B), Relative phylogenetic diversity (RPD). (C), Functional diversity (FD). (D), Relative functional diversity (RFD). Trendlines in (a, c) fit with general additive model using `r func("geom_smooth")` function in the `r pack("ggplot2")` R package [@Wickham2016].

@@

**`r s_figure("endemism-restricted")`**. Phylogenetic endemism of the ferns of Japan measured using CANAPE (categorical analysis of neo- and paleo-endemism), restricted dataset including only taxa endemic to Japan. 'paleo' indicates paleoendemic areas (those with an overabundance of long branches); 'neo' indicates neodendemic areas (those with an overabundance of short branches); 'mixed' indicates significantly endemic sites that have neither an overabundance of short nor long branches; 'super' indicates highly significantly endemic sites that have neither an overabundance of short nor long branches. For details, see Methods.

@@

**`r s_figure("k-plot")`**. Selection of *k* for bioregions. (A), taxonomic bioregions. (B), phylogenetic bioregions. Color indicates optimal *k* selected selected by the `r func("optimal_phyloregion")` function in the R package `r pack("phyloregion")` (black, non-optimal; red, optimal).

@@

# Map taxonomic clusters to grid cells
clusters_tax <-
  biodiv_ferns_spatial %>%
  as_tibble() %>%
  select(grids, cluster = taxonomic_cluster)

cluster_nodes_tax_tbl <- map_df(unique(clusters_tax$cluster), ~map_cluster_to_nodes(dendro_tax, clusters_tax, .)) %>%
  # Lump regions with fewer grid-cells than cutoff
  mutate(cluster = as.factor(cluster) %>% fct_lump_min(bioregion_cutoff))

# Make sure cutoff matches caption
assert_that(bioregion_cutoff == 2, msg = "Bioregions cutoff for lumping not 2!")

@@

**`r s_figure("dendrogram")`**. Bioregion dendrograms. Tips are sites (names not shown). Dendrograms generated by clustering taxonomic (Sørensen) or phylogenetic (PhyloSor) distances between sites (see Materials and Methods). Color shows bioregion; bioregions not consisting of more than `r english2(bioregion_cutoff)` grid-cells each are lumped into the "Other" category. (A), taxonomic bioregions. (B), phylogenetic bioregions. Dendrograms visualized with the `r pack("ggtree")` R package [@Yu2017].
\clearpage

@@

# Map phylogenetic clusters to grid cells
clusters_phy <-
  biodiv_ferns_spatial %>%
  as_tibble() %>%
  select(grids, cluster = phylo_cluster)

cluster_nodes_phy_tbl <- map_df(unique(clusters_phy$cluster), ~map_cluster_to_nodes(dendro_phy, clusters_phy, .)) %>%
  # Lump regions with fewer grid-cells than cutoff
  mutate(cluster = as.factor(cluster) %>% fct_lump_min(bioregion_cutoff))

@@

**`r s_figure("genus-map")`**. Map of species richness for selected genera of ferns of Japan. (A), *Equisetum* is an example of a possible refugial lineage with greater species richness in northern Japan. (B), *Deparia* is an example of a genus with possible recent radiation in southern Honshu. (C)--(F) show genera with high rates of apomictic taxa (`r s_figure("apo-genus")`), which tend to be greatest in southern Honshu.

>>>>

```{r load-data-si, cache = FALSE}
tar_load(c(
  "japan_shp",
  "inext_res",
  "redundancy_by_res",
  "redundancy_summary",
  "biodiv_ferns_spatial",
  "biodiv_ferns_endemic_spatial",
  "k_taxonomy",
  "k_phylogeny",
  "traits_nmds",
  "ppgi",
  "phy_sig_results",
  "binary_sig_results",
  "lrt_results",
  "japan_shp",
  "morans_i",
  "t_test_results",
  "aic_env_repro",
  "comm_ferns",
  "repro_data",
  "japan_pterido_tree",
  "model_fits_not_ult",
  "biodiv_ferns_cent",
  "protected_areas",
  "deer_range",
  "biodiv_ferns_spatial_lumped",
  "bioregion_cutoff",
  "bioregions_all_labels",
  "ms_functions"
))
```

@@

**`r s_figure("grain-size")`**. Effect of grain size on sampling redundancy. (A) Percent of grid cells with redundancy > 0.5 by grain size (side length of square grid cells set to 10 km, 20 km, 30 km, or 40 km). (B) Percent of grid cells with redundancy = 0 by grain size. (C) Distribution of redundancy values by grain size.

@@

**`r s_figure("envir")`**. Reproductive mode and environmental data in 20 km × 20 km grid cells across Japan. (A) rolling mean area (km^2^) in 20 km latitudinal bands. (B) % apomictic taxa. (C) annual precipitation. (D) percipitation seasonality (coefficient of variation). (E) mean annual temperature. (F) temperature seasonality (standard deviation × 100). Data excluded from models not shown in this and subsequent plots unless otherwise mentioned.

@@

**`r s_figure("abundance")`**. Observed number of specimens (A) and sampling redundancy (B) per 20 km × 20 km grid cell in the ferns of Japan.

@@

**`r s_figure("raw-lat")`**. (A) Raw taxonomic diversity, (B) phylogenetic diversity, and (C) functional diversity of the ferns of Japan plotted by latitude. For (B) and (C), raw branchlengths were transformed to relative values by dividing each branchlength by the sum of all branchlengths. Trendlines fit with general additive model using `r func("geom_smooth")` function in the `r pack("ggplot2")` R package [@Wickham2016].

@@

**`r s_figure("raw-biplot")`**. Relationships between observed phylogenetic and functional diversity and taxonomic richness in the ferns of Japan. (A) Phylogenetic diversity (PD). (B) Relative phylogenetic diversity (RPD). (C) Functional diversity (FD). (D) Relative functional diversity (RFD). Trendlines in (A, C) fit with general additive model using `r func("geom_smooth")` function in the `r pack("ggplot2")` R package [@Wickham2016].

@@

**`r s_figure("endemism-restricted")`**. Phylogenetic endemism of the ferns of Japan measured using CANAPE (categorical analysis of neo- and paleo-endemism), restricted dataset including only taxa endemic to Japan. 'neo' indicates neodendemic areas (those with an overabundance of short branches); 'paleo' indicates paleoendemic areas (those with an overabundance of long branches); 'mixed' indicates significantly endemic sites that have neither an overabundance of short nor long branches; 'super' indicates highly significantly endemic sites that have neither an overabundance of short nor long branches. For details, see Materials and Methods.

@@

**`r s_figure("k-plot")`**. Selection of *k* for bioregions. (A) taxonomic bioregions. (B) phylogenetic bioregions. Color indicates optimal *k* selected selected by the `r func("optimal_phyloregion")` function in the R package `r pack("phyloregion")` (black, non-optimal; red, optimal).

@@

# Map taxonomic clusters to grid cells. Include original ("old") clusters
# so that grouping of "other" works correctly even if not monophyletic
clusters_tax <-
bioregions_all_labels %>%
  # Filter to grid-cells used in modeling
  inner_join(
    select(biodiv_ferns_spatial, grids),
    by = "grids"
  ) %>%
  add_count(old_taxonomic_cluster) %>%
  mutate(
    taxonomic_cluster = case_when(
      n <= bioregion_cutoff ~ paste(taxonomic_cluster, old_taxonomic_cluster, sep = "_"),
      TRUE ~ taxonomic_cluster
      )
  ) %>%
  select(-n) %>%
  select(grids, cluster = taxonomic_cluster) %>%
  arrange(grids)

cluster_nodes_tax_tbl <- map_df(unique(clusters_tax$cluster), ~map_cluster_to_nodes(dendro_tax, clusters_tax, .)) %>%
  # Lump regions with fewer grid-cells than cutoff
  mutate(cluster = as.factor(cluster) %>% fct_lump_min(bioregion_cutoff + 1))

@@

**`r s_figure("dendrogram")`**. Bioregion dendrograms. Tips are sites (names not shown). Dendrograms generated by clustering taxonomic (Sørensen) or phylogenetic (PhyloSor) distances between sites (see Materials and Methods). Color shows bioregion; bioregions not consisting of more than `r english2(bioregion_cutoff)` grid-cells each are lumped into the "Other" category. (A) taxonomic bioregions. (B) phylogenetic bioregions. Dendrograms visualized with the `r pack("ggtree")` R package [@Yu2017].
\clearpage

@@

# Map phylogenetic clusters to grid cells. Include original ("old") clusters
# so that grouping of "other" works correctly even if not monophyletic
clusters_phy <-
bioregions_all_labels %>%
  # Filter to grid-cells used in modeling
  inner_join(
    select(biodiv_ferns_spatial, grids),
    by = "grids"
  ) %>%
  add_count(old_phylo_cluster) %>%
  mutate(
    phylo_cluster = case_when(
      n <= bioregion_cutoff ~ paste(phylo_cluster, old_phylo_cluster, sep = "_"),
      TRUE ~ phylo_cluster
      )
  ) %>%
  select(-n) %>%
  select(grids, cluster = phylo_cluster) %>%
  arrange(grids)

cluster_nodes_phy_tbl <- map_df(unique(clusters_phy$cluster), ~map_cluster_to_nodes(dendro_phy, clusters_phy, .)) %>%
  # Lump regions with fewer grid-cells than cutoff
  mutate(cluster = as.factor(cluster) %>% fct_lump_min(bioregion_cutoff + 1))

@@

**`r s_figure("genus-map")`**. Map of species richness for selected genera of ferns of Japan. (A) *Equisetum* is an example of a possible refugial lineage with greater species richness in northern Japan. (B) *Deparia* is an example of a genus with possible recent radiation in southern Honshu. (C)--(F) show genera with high rates of apomictic taxa (`r s_figure("apo-genus")`), which tend to be greatest in southern Honshu.

### 2022-01-21 ms/manuscript.Rmd
Update MS
Fix dendrogram in SI where non-monophyletic 'other' was causing large portion of tree to be colored grey

**Methods:** We compiled a community dataset of `r nrow(comm_ferns) %>% scales::number(big.mark = ",")` 20 km × 20 km grid-cells including `r ncol(comm_ferns)` taxa based on > `r occ_sum$filtered$n_specimens$value %>% plyr::round_any(1000, f = floor) %>% number(big.mark = ",")` specimen records.
We combined this with a phylogeny and functional traits to analyze taxonomic, phylogenetic, and functional diversity, and modeled biodiversity metrics in response to environmental factors and reproductive mode.
Hierarchical clustering was used to analyze bioregions.
Conservation status and threats were assessed by comparing the overlap of significantly diverse grid-cells with conservation zones and range maps of native Japanese deer.

**Results:** Taxonomic richness was highest at mid-latitudes.
Phylogenetic and functional diversity and phylogenetic endemism were highest in small southern islands.
Relative phylogenetic and functional diversity were high at both high and low latitudes, and low at mid-latitudes.
Grid-cells were grouped in to `r maj_cluster_count$phylo_cluster$n %>% english2` (phylogenetic) or `r maj_cluster_count$taxonomic_cluster$n %>% english2` (taxonomic) major bioregions.
Temperature and apomixis were identified as drivers of taxonomic and phylogenetic diversity.
Conservation status was generally high for grid-cells with significantly high biodiversity, but the threat due to herbivory was greater for taxonomic richness than other metrics.

**Conclusions:** The integrative approach used here reveals previously undetected patterns and drivers of biodiversity in the ferns of Japan.
Future conservation efforts should recognize that threats can vary by biodiversity metric and consider multiple multiple metrics when establishing conservation priorities.

@@

However, a comprehensive understanding of the relationships between richness and other metrics requires densely sampled data, and such datasets are rare on the regional (country) scale.
The ferns of Japan are excellent model system because they have been the target of intense botanical interest for several decades and are densely sampled [reviewed in @Ebihara2019b].
The ferns of Japan include `r total_fern_taxa` native, non-hybrid taxa (including species and varieties) and hundreds of hybrids [@Ebihara2019b].
The availability of detailed distribution data [distribution maps at the ca. 10 km scale for nearly all species\; @Ebihara2016b; @Ebihara2017], trait data [multiple quantitative and qualitative traits compiled for identification of nearly all species\; @Ebihara2019b], and DNA sequences for > 97% of species [@Ebihara2010; @Ebihara2019b\; all coverage statistics exclude hybrids] make the ferns of Japan an ideal system for investigating the relationships between, and drivers of, multiple dimensions of biodiversity.

@@

We used a list of native, non-hybrid fern specimens deposited at herbaria in Japan to quantify occurrences [@Ebihara2016b; @Ebihara2017; @Ebihara2019b].
We chose to use this over other sources (`r eg`, GBIF) because of its high quality, which obviates many cleaning steps otherwise needed when working with large, publicly available datasets.
Furthermore, the overall sampling completeness of this dataset is also high (`r s_figure("sampling-curve")` in Appendix S1; see Supplemental Data with this article).
The original list comprises `r occ_sum$unfiltered$n_specimens$value %>% scales::number(big.mark = ",")` specimens representing `r occ_sum$unfiltered$n_taxa$value` terminal taxa at the rank of species, subspecies, form, or variety, excluding non-native taxa and nothotaxa.
For the purposes of analysis, we treated infraspecific taxa as distinct species and hereafter use "taxon" and "species" interchangeably to refer to entities at this rank, unless otherwise indicated.
All names are standardized to a common taxonomy, the Japan Green List v. 1.01 (http://www.rdplants.org/gl/).

Occurrence records were georeferenced and thoroughly vetted as follows [@Ebihara2016b; @Ebihara2017; @Ebihara2019b].
For specimens that lacked latitude and longitude data, georeferencing was done by mapping collection site names to a set of standard ca. 10 × 10 km grid squares defined by the Statistics Bureau of Japan (the "second-degree mesh"; http://www.stat.go.jp/english/data/mesh/05.html), and the centroid of the second-degree mesh cell used as the specimen location.
In the case that the collection site could not be mapped to a single second-degree mesh cell, it was excluded.
Next, occurrence maps were generated for each taxon showing presence or absence in each second-degree mesh cell.
In the case that a given taxon appeared insufficiently sampled (`r ie`, not present on the map where it would typically be expected to occur), AE and members of the Nippon Fernist Club (local botanists familiar with the ferns of Japan) searched for additional specimens, which were then added to the list.
The occurrence maps were iteratively refined until the vast majority of known Japanese ferns had been observed across their expected ranges; the resulting set of occurrence records can be considered accurate to ca. 10 km (the grain size of the second-degree mesh map).

@@

Sequences for the plastid, coding *rbcL* gene are available for ca. 97% of the Japanese fern flora [@Ebihara2019b]; these sequences originate from samples that were identified using the same taxonomic system as the occurrence data [@Ebihara2010; @Ebihara2019b], so we used these preferentially over other data available on GenBank that might include misidentifications or different taxon concepts.
However, using this dataset alone for phylogenetic analysis suffers two drawbacks: it is not densely sampled enough to resolve some internal nodes, and it lacks many lineages that have fossils available for molecular dating.
Furthermore, use of community phylogenies generated by only by sampling the species present in the local community has been shown to produce spurious results in simulation studies [@Park2018].
Because one goal of the study is to understand the distribution of paleo- vs. neo-endemic areas (defined in units of time), we require an ultrametric tree.
Therefore, to obtain a robust, ultrametric tree, we combined the *rbcL* data of @Ebihara2019b (`r n_japan_fern_taxa` taxa) with a globally sampled plastid dataset including all fern sequences on GenBank for four widely-sequenced plastid genes (*atpA*, *atpB*, *rbcL*, and *rps4*; `r n_ftol_fern_taxa` taxa) and `r n_plastome_genes` other coding, single-copy plastid genes extracted from `r n_plastome_fern_taxa` complete fern plastomes (Nitta et al., in prep.) (`r ape::Ntip(plastome_tree) %>% scales::number(big.mark = ",")` OTUs total, including outgroup taxa).
This gene sampling is comparable to a recent global fern phylogeny that resolved relationships across ca. 4,000 taxa using six plastid markers [@Testo2016a], and the addition of plastome data can be expected to increase support along the backbone.
We conducted maximum-likelihood phylogenetic analysis with `r software("IQ-TREE")` v1.6.12 [@Nguyen2015] under the GTR+I+G model (all sites concatenated into a single data matrix), and assessed node support with 1,000 ultrafast bootstrap replicates [@Minh2013].
Molecular dating was conducted with `r software("treePL")` v1.0 [@Smith2012] using `r n_fossil_points` fossil calibration points after @Testo2016a, with the exception of treating fossil *Kuylisporites* as belonging to crown *Cyathea*, not *Cyathea* + *Alsophila* [@Loiseau2020].
The root of the tree (most recent common ancestor of all land plants) was fixed to 475 mya [@Testo2016a].
We trimmed the resulting ultrametric tree to only Japanese taxa.
`r software("treePL")` requires branch lengths to be > 0, and it sets extremely small branches to an arbitrary minimum positive length.
This may artificially add branch length between taxa with identical DNA sequences, which occur multiple times in the Japanese fern *rbcL* dataset.
We converted any clades in the dated tree consisting of identical sequences to polytomies with the node at 0 mya (`r ie`, the present).
This tree was used for all subsequent analysis unless indicated otherwise.

@@

One element of biodiversity that we did not consider here is hybrids (nothotaxa).
The Japanese fern flora includes `r n_hybrid_fern_taxa` nothotaxa [@Ebihara2019b].
Although hybrids are often thought of as evolutionary "dead-ends", in some cases they are able to interbreed with diploids and thus contribute to diversification [@Barrington1989].
Future studies should consider the distribution of hybrids in Japan and their possible effects on biodiversity.
Methods for calculating PD would need to be modified to take into account multiple lineages contributing to one terminal taxon.

@@

The authors thank Members of the Iwasaki Lab at the University of Tokyo and two anonymous reviewers for providing comments that improved the manuscript.
Shawn Laffan provided advice about CANAPE.
This study was supported by JSPS KAKENHI Grant Number 16H06279.

@@

**`r figure("endemism")`**. Phylogenetic endemism of the ferns of Japan measured using CANAPE (categorical analysis of neo- and paleo-endemism). 'paleo' indicates areas with an overabundance of rare long branches; 'neo' indicates areas with an overabundance of rare short branches; 'mixed' indicates significantly endemic sites that have neither an overabundance of short nor long branches; 'super' indicates highly significantly endemic sites that have neither an overabundance of short nor long branches. For details, see Materials and Methods.

>>>>

**Methods:** We compiled a community dataset of `r nrow(comm_ferns) %>% scales::number(big.mark = ",")` 20 km × 20 km grid-cells including `r ncol(comm_ferns)` taxa based on > `r occ_sum$filtered$n_specimens$value %>% plyr::round_any(1000, f = floor) %>% number(big.mark = ",")` specimen records.
We combined this with a phylogeny and functional traits to analyze taxonomic, phylogenetic, and functional diversity, and modeled biodiversity metrics in response to environmental factors and reproductive mode.
Hierarchical clustering was used to delimit bioregions.
Conservation status and threats were assessed by comparing the overlap of significantly diverse grid-cells with conservation zones and range maps of native Japanese deer.

**Results:** Taxonomic richness was highest at mid-latitudes.
Phylogenetic and functional diversity and phylogenetic endemism were highest in small southern islands.
Relative phylogenetic and functional diversity were high at high and low latitudes, and low at mid-latitudes.
Grid-cells were grouped into `r maj_cluster_count$phylo_cluster$n %>% english2` (phylogenetic) or `r maj_cluster_count$taxonomic_cluster$n %>% english2` (taxonomic) major bioregions.
Temperature and apomixis were identified as drivers of biodiversity patterns.
Conservation status was generally high for grid-cells with significantly high biodiversity, but the threat due to herbivory by deer was greater for taxonomic richness than other metrics.

**Conclusions:** Our integrative approach reveals previously undetected patterns and drivers of biodiversity in the ferns of Japan.
Future conservation efforts should recognize that threats can vary by biodiversity metric and consider multiple multiple metrics when establishing conservation priorities.

@@

However, a comprehensive understanding of the relationships between richness and other metrics requires densely sampled data, and such datasets are rare on the regional (country) scale.
The ferns of Japan are excellent model system because they have been the target of intense botanical interest for several decades and are densely sampled [reviewed in @Ebihara2019b].
The ferns of Japan include `r total_fern_taxa` native, non-hybrid taxa (including species and varieties) and hundreds of hybrids [nothotaxa\; @Ebihara2019b].
The availability of detailed distribution data [distribution maps at the ca. 10 km scale for nearly all species\; @Ebihara2016b; @Ebihara2017], trait data [multiple quantitative and qualitative traits compiled for identification of nearly all species\; @Ebihara2019b], and DNA sequences for > 97% of species [@Ebihara2010; @Ebihara2019b\; all coverage statistics exclude nothotaxa] make the ferns of Japan an ideal system for investigating the relationships between, and drivers of, multiple dimensions of biodiversity.

@@

We used a list of native, non-hybrid fern specimens deposited at herbaria in Japan to quantify occurrences [@Ebihara2016b; @Ebihara2017; @Ebihara2019b].
We chose to use this over other sources such as the Global Biodiversity Information Facility (GBIF) because of its high quality, which obviates many cleaning steps otherwise needed when working with large, publicly available datasets.
Furthermore, the overall sampling completeness of this dataset is also high (`r s_figure("sampling-curve")` in Appendix S1; see Supplemental Data with this article).
The original list comprises `r occ_sum$unfiltered$n_specimens$value %>% scales::number(big.mark = ",")` specimens representing `r occ_sum$unfiltered$n_taxa$value` terminal taxa at the rank of species, subspecies, form, or variety, excluding non-native taxa and nothotaxa.
We treated infraspecific taxa as distinct species during analysis and hereafter use "taxon" and "species" interchangeably to refer to entities at this rank, unless otherwise indicated.
All names are standardized to a common taxonomy, the Japan Green List v. 1.01 (http://www.rdplants.org/gl/).

Occurrence records were georeferenced and thoroughly vetted as follows [@Ebihara2016b; @Ebihara2017; @Ebihara2019b].
For specimens that lacked latitude and longitude data, georeferencing was done by mapping collection site names to a set of standard ca. 10 km × 10 km grid squares defined by the Statistics Bureau of Japan (the "second-degree mesh"; http://www.stat.go.jp/english/data/mesh/05.html), and the centroid of the second-degree mesh cell used as the specimen location.
In the case that the collection site could not be mapped to a single second-degree mesh cell, it was excluded.
Next, occurrence maps were generated for each taxon showing presence or absence in each second-degree mesh cell.
In the case that a given taxon appeared insufficiently sampled (`r ie`, not present on the map where it would typically be expected to occur), AE and members of the Nippon Fernist Club (local botanists familiar with the ferns of Japan) searched for additional specimens, which were then added to the list.
The occurrence maps were iteratively refined until the vast majority of known Japanese ferns had been observed across their expected ranges; the resulting set of occurrence records can be considered accurate to ca. 10 km (the grain size of the second-degree mesh map).

@@

Sequences for the plastid, coding *rbcL* gene are available for ca. 97% of the Japanese fern flora [@Ebihara2019b]; these sequences originate from samples that were identified using the same taxonomic system as the occurrence data [@Ebihara2010; @Ebihara2019b], so we used these preferentially over other data available on GenBank that might include misidentifications or different taxon concepts.
However, using this dataset alone for phylogenetic analysis suffers two drawbacks: it is not densely sampled enough to resolve some internal nodes, and it lacks many lineages that have fossils available for molecular dating.
Furthermore, use of community phylogenies generated by only by sampling the species present in the local community has been shown to produce spurious results in simulation studies [@Park2018].
Because one goal of the study is to understand the distribution of paleo- vs. neo-endemic areas (defined in units of time), we require an ultrametric tree.
Therefore, to obtain a robust, ultrametric tree, we combined the *rbcL* data of @Ebihara2019b (`r n_japan_fern_taxa` taxa) with a globally sampled plastid dataset including all fern sequences on GenBank for four widely-sequenced plastid genes (*atpA*, *atpB*, *rbcL*, and *rps4*; `r n_ftol_fern_taxa` taxa) and `r n_plastome_genes` other coding, single-copy plastid genes extracted from `r n_plastome_fern_taxa` complete fern plastomes (Nitta et al., in prep.) (`r ape::Ntip(plastome_tree) %>% scales::number(big.mark = ",")` OTUs total, including outgroup taxa).
This gene sampling is comparable to a recent global fern phylogeny that resolved relationships across ca. 4,000 taxa using six plastid markers [@Testo2016a], and the addition of plastome data can be expected to increase support along the backbone.

We conducted maximum-likelihood phylogenetic analysis with `r software("IQ-TREE")` v1.6.12 [@Nguyen2015] under the GTR+I+G model (all sites concatenated into a single data matrix), and assessed node support with 1,000 ultrafast bootstrap replicates [@Minh2013].
Molecular dating was conducted with `r software("treePL")` v1.0 [@Smith2012] using `r n_fossil_points` fossil calibration points after @Testo2016a, with the exception of treating fossil *Kuylisporites* as belonging to crown *Cyathea*, not *Cyathea* + *Alsophila* [@Loiseau2020].
The root of the tree (most recent common ancestor of all land plants) was fixed to 475 mya [@Testo2016a].
We trimmed the resulting ultrametric tree to only Japanese taxa.
`r software("treePL")` requires branch lengths to be > 0, and it sets extremely small branches to an arbitrary minimum positive length.
This may artificially add branch length between taxa with identical DNA sequences, which occur multiple times in the Japanese fern *rbcL* dataset.
We converted any clades in the dated tree consisting of identical sequences to polytomies with the node at 0 mya (`r ie`, the present).
This tree was used for all subsequent analysis unless indicated otherwise.

@@

One element of biodiversity that we did not consider here is hybrids.
The Japanese fern flora includes `r n_hybrid_fern_taxa` nothotaxa [@Ebihara2019b].
Although hybrids are often thought of as evolutionary "dead-ends", in some cases they are able to interbreed with diploids and thus contribute to diversification [@Barrington1989].
Future studies should consider the distribution of hybrids in Japan and their possible effects on biodiversity.
Methods for calculating PD would need to be modified to take into account multiple lineages contributing to one terminal taxon.

@@

The authors thank Members of the Iwasaki Lab at the University of Tokyo and two anonymous reviewers for providing comments that improved the manuscript.
Shawn Laffan provided advice about CANAPE.
This study was supported by JSPS KAKENHI Grant Number 16H06279.
The authors are grateful to members of the Nippon Fernist Club for their cooperation and efforts towards quantifying and conserving the biodiversity of the ferns of Japan.

@@

**`r figure("endemism")`**. Phylogenetic endemism of the ferns of Japan measured using CANAPE (categorical analysis of neo- and paleo-endemism). 'neo' indicates areas with an overabundance of rare short branches; 'paleo' indicates areas with an overabundance of rare long branches; 'mixed' indicates significantly endemic sites that have neither an overabundance of short nor long branches; 'super' indicates highly significantly endemic sites that have neither an overabundance of short nor long branches. For details, see Materials and Methods.
